---
title: "Project Assignment 1 (RepData)"
author: "guo windy"
date: "Saturday, December 13, 2014"
output: html_document



---


**What is mean total number of steps taken per day?**  
1.  Make a histogram of the total number of steps taken each day  
2.  Calculate and report the mean and median total number of steps taken per day  

```{r}
## loading data
data<-read.csv("activity.csv")
## sum of the steps in each date (ignore the missing values in the dataset)
sumsteps<-tapply(data[,1], data$date, sum)
## plot the histogram
hist(sumsteps,main="Histogram plot of the steps in each date", xlab="Sum of steps per day")
## Calculate and report the mean and median 
mean_sumsteps<-mean(sumsteps,na.rm=T)
median_sumsteps<-median(sumsteps, na.rm=T)
print(mean_sumsteps)
print(median_sumsteps)
```
* ----The mean value of the total number of steps taken per day is 10766.2.
* ----The median value of the total number of steps taken per day is 10765.

**What is the average daily activity pattern?**    
1.  Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)    
2.	Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?   

```{r}
## calculate the mean step based on the interval over all the days in the dataset
s5m<-aggregate(x = data$steps, by = list(data$interval), FUN = "mean", na.rm=TRUE)
## give appropriate names
colnames(s5m) <- c("interval","mean_steps")

## spot out the maximum value
maxp<-which.max(s5m$mean_steps)
maxdatapoint<-s5m[maxp ,]
print(maxdatapoint)
```
* ---- The "835" interval is the maximum one!

```{r}
## give the plot
library("ggplot2")
ggplot(data=s5m, aes(x=interval, y=mean_steps))+ 
        geom_line(size =1,colour="#000099") +
        ggtitle(" The average number of steps taken at 5-minute interval")+ 
        ylab("mean of steps") +
		## mark the maximum point with red circle 
        geom_point(aes( x=835,y=206.169), color="red",shape = 1,size = 4) 
```


**Imputing missing values**    

1.      Calculate and report the total number of missing values in the dataset (i.e. the total number of rows withNAs)   
2.	Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.    
3.	Create a new dataset that is equal to the original dataset but with the missing data filled in.   
4.	Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?     

```{r}
##  the total number of the missing values
numofna<-length(which(is.na(data$steps)))
print(numofna)
```
* ---- Total number of NA is 2304.
```{r}
##  strategy A: fill with the mean/median for that day
##  calculate and construct the data set "sd" with the values of the mean of steps for each date
sd<-aggregate(x = data$steps, by = list(data$date), FUN = "mean", na.rm=TRUE)

## find the NA in the data set "sd"
mean_na<-which(is.na(sd[,2]))
## calculate and replace NA with the mean step of whole data set
mean_all<-mean(sd[,2],na.rm=TRUE) ## the mean step of the whole data set
sd[mean_na,2]<-mean_all           ## replacement

## Now "sd" is a data set without missing value
## we can fill in them with the mean of that day data

## first find the index of the missing values
naindex<-which(is.na(data$steps))
## their corresponding "date"
nadate<-data[naindex,2];cha_nadate<-as.character(nadate)

## newdata set 
newdata<-data
## filling now
for (i in 1:length(cha_nadate)) {        
       dateindex<- which(sd[,1]==cha_nadate[i])
       newdata[naindex[i],1]<- sd[dateindex,2]        
}

## sum of the steps in each date (ignore the missing values in the dataset)
new_sumsteps<-tapply(newdata[,1], newdata$date, sum);
## plot the histogram
hist(new_sumsteps,main="Histogram plot of the steps in each date in new data set", xlab="Sum of steps per day",col="red")
## Calculate and report the mean and median 
mean_new_sumsteps<-mean(new_sumsteps)
median_new_sumsteps<-median(new_sumsteps)
```
Do these values differ from the estimates from the first part of the assignment?

```{r}
## the mean and median of these two different data(raw data and filled data)
c(mean_sumsteps, mean_new_sumsteps)
c(median_sumsteps,median_new_sumsteps)
```
* ---- From the comparison of the above values, we can see the minor difference. 

What is the impact of imputing missing data on the estimates of the total daily number of steps?   

* ----since we filled in NA with the mean value, so total mean will not change, but the median will slightly change.



**Are there differences in activity patterns between weekdays and weekends?**   
1.	Create a new factor variable in the dataset with two levels ? ?weekday? and ?weekend? indicating whether a given date is a weekday or weekend day.   
2.	Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.   

```{r}
## add a new column with weekdays()
newdata$week_day<-weekdays(as.Date(newdata$date))
##  change them into two levels 
newdata[newdata$week_day=="Sunday",4]<-"weekend"
newdata[newdata$week_day=="Satday",4]<-"weekend"
newdata[newdata$week_day!="weekend",4]<-"weekday"

## calculate the mean steps of 5 mins separated for weekend and weekday
weeks5m<-aggregate(steps~interval+week_day,newdata, FUN = "mean")
## plot in the panel mode
library("lattice")
colnames(weeks5m) <- c("interval","week_day","mean_steps")
xyplot(mean_steps ~ interval | week_day, 
       data=weeks5m,layout = c(1,2),type = "a",
       ylab = "Number of steps")

```